---
title: "Spatial separable model"
output: 
  html_notebook: 
    toc: yes

---

This R code is for smoothing positivity rate since its raw data contains lots of 0 and 1. 

## Data process

```{r}
suppressPackageStartupMessages({
library(stringi); library(dplyr);library(tidyverse); library(tigris); library(tidycensus); library(tmap); library(readxl); library(INLA); library(spdep);library(ggcorrplot);library(rgdal);library(ggmap);library(ggplot2)})

setwd("/Users/Ana/Downloads/Testing_gap/spatial_temporal_model/")
load("dat.Rdata")

```

### Testing data

#### Read in COVID19 testing data
```{r}
library("DClusterm")
library(yarrr)
#### read in raw gap data###
data_block_gap <- read.csv("processed_data.csv")
data_block_gap$block_grp <- as.character(data_block_gap$block_grp)

### set the new test frequency to 1, if the positive_freq is NA
data_block_gap$positive_freq[data_block_gap$test_freq == 0] <- NA
data_block_gap$new_test_freq <- data_block_gap$test_freq 
data_block_gap$new_test_freq[data_block_gap$test_freq == 0] <- 1 #to make prediction change 0 to 1
#View(cbind(new_test_freq, data_block_gap$test_freq))

```


```{r}
data_block_gap %>% dplyr::select(block_grp, week, test_freq, positive_freq, census_pop) %>% arrange(desc(test_freq)) 
```

### Spatial data

read in the spatial info and merge CBGs where population is 0 with corresponding neighbor.
```{r}
#Need to merge tract 980001 to 010203, 980100 to 013206, 990000 to 012305
#Make the merged tracts have the same ID
cameron$ID[cameron$TRACTCE=="980100"] <- 225
cameron$ID[cameron$TRACTCE=="980001"] <- 21
cameron$ID[cameron$TRACTCE=="990000"] <- 58


library(maptools)
gpclibPermit()
cameron2 <- unionSpatialPolygons(cameron, cameron$ID)

 # plot(cameron2)
 # plot(cameron)

cameron.df <- as(cameron, "data.frame")
cameron.df <- subset(cameron.df, ! cameron.df$TRACTCE %in% c ("980100","980001","990000"))
row.names(cameron.df) <- as.character(cameron.df$ID)

# Reconvert data frame to SpatialPolygons
cameron <- SpatialPolygonsDataFrame(cameron2, cameron.df)
Cameron<- geocode("Cameron county, TX")  #lon -97.5  lat 26.1
Cameron_center <- c(-97.5, 26.1)

## remove the columns that do not need, and write the shapefile for the selected areas 
cameron$ID <- NULL
cameron$FUNCSTAT <- NULL
#writeOGR(cameron, dsn="Cameron_CBG_222", layer="Cameron_CBG_222", driver="ESRI Shapefile")
cameron222 <-  readOGR(dsn='Cameron_CBG_222', 'Cameron_CBG_222', verbose=F)
cameron222$ID <- seq_along(cameron222[,1]) 

### tricks to use for sanity check
### to find the centroids of each county
centroids <- coordinates(cameron222)
### add county names
plot(cameron222)
text(centroids, labels=cameron222$GEOID, cex=0.8, col="royalblue") ### not as good as the interactive but can be used for quick and dirty check

```

Generate adjacency matrix
```{r}
dat.adj <- poly2nb(cameron222)
dat.mat <- as(nb2mat(dat.adj, style = "B"), "Matrix")

```

#### Merge testing data with spatial info
```{r}
#merge gap data set with spatial info
data_with_sp <- left_join(data_block_gap, cameron222@data , by = c("block_grp" = "GEOID")) 

dat2 <- data_with_sp %>% 
  dplyr::select(block_grp, week, test_freq, positive_freq, census_pop, ID) 
map(dat2, summary)
```

This dataset is already coded such that if the number of tests are 0, the number of positive cases are NA. This is already done, and there are a total of 651 NAs. The question now is to deal with the offset part. 

```{r}
summary(data_with_sp$positive_freq)
```


## Models - no interaction

### Model with logit link (default)
```{r}
f1  <- positive_freq ~ 1 + f(week, model = "rw1") + f(ID, model = "bym", graph = dat.mat) 

## model with logit link 
model_logit <-inla(f1, data = data_with_sp,  family = "binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),
                control.predictor = list(compute = TRUE, link=1))
summary(model_logit)
fit1_logit <- model_logit$summary.fitted.value[, 1]
summary(fit1_logit)

```

### Model with probit link
```{r}
model_probit <-inla(f1, data = data_with_sp,  family = "binomial", Ntrials= new_test_freq, control.family = list(link = "probit"),
                control.predictor = list(compute = TRUE, link=1))
summary(model_probit)
fit1_probit <- model_probit$summary.fitted.value[, 1]
summary(fit1_probit)
```

### Model with cloglog link
```{r}
### cloglog
model_cloglog <-inla(f1, data = data_with_sp,  family = "binomial", Ntrials= new_test_freq, control.family = list(link = "cloglog"),
                control.predictor = list(compute = TRUE, link=1))
summary(model_cloglog)
fit1_cloglog <- model_cloglog$summary.fitted.value[, 1]
summary(fit1_cloglog)
```


### Comparison between various outputs
```{r}
df <- data_with_sp %>% mutate(rate_posi = positive_freq/test_freq)
df$rate_logit <- fit1_logit
df$rate_probit <- fit1_probit
df$rate_cloglog <- fit1_cloglog
head(df, n=10)
```

Now make a plot to compare the estimates 
```{r}
tmp <- df %>% pivot_longer(cols=starts_with("rate"), names_to="Mod", values_to="Est")
head(tmp, n=5)
ggplot(tmp, aes(week, Est)) + geom_point(aes(colour = factor(Mod)), size = 1)
```

 
Use the simple scatterplot with points color-coded by model. So link functions do not make a lot of different, but compared to the observed there was substantial shrinkage.
```{r}
plot(df$rate_logit, df$rate_posi, col=2, xlab="Logit Est", ylab="Est", xlim=c(0, 1), ylim=c(0,1))
abline(a=0, b=1, col="grey", lwd=2)

points(df$rate_logit, df$rate_cloglog, col=3)
points(df$rate_logit, df$rate_probit, col=4)
legend("bottomright", c("observed", "cloglog", "probit"), pch=1, col=c(2, 3, 4), bty="n")
```


```{r}
plot(df$rate_posi, df$rate_logit, col=2, xlab="Observed Positivity Rate", ylab="Logit Est", xlim=c(0, 1), ylim=c(0,1))
abline(a=0, b=1, col="grey", lwd=2)
```
## Models - with interaction

The fit from this simple spatiotemporal main effects model doesn't appear to be that good. We can try the space-time interaction models to have "less smoothing". These models can be time consuming to run!!!

### Models using rw2 latend effect
```{r}
dat3 <- data_with_sp %>% mutate(ID.area = ID, ID.time=week, ID.time2=week, ID.area.time = 1:nrow(data_with_sp))
#--- Type I interaction ---#
formula.intI <- positive_freq ~ f(ID.area, model="bym", graph=dat.mat) + f(ID.time, model="rw2") + 
  f(ID.time2,model="iid") + f(ID.area.time, model="iid")
mod.intI <- inla(formula.intI, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type II interaction ---#
dat3$ID.area.int <- dat3$ID.area
dat3$ID.time.int <- dat3$ID.time

formula.intII <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw2") + f(ID.time2,model="iid") +
  f(ID.area.int, model="iid", group=ID.time.int, control.group=list(model="rw2")) 
mod.intII <- inla(formula.intII, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type III interaction ---#
formula.intIII<- positive_freq ~  f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw2") + f(ID.time2,model="iid") +
  f(ID.time.int,model="iid", group=ID.area.int, control.group=list(model="besag", graph=dat.mat))
mod.intIII <- inla(formula.intIII, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type IV interaction ---#
formula.intIV <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw2") + f(ID.time2,model="iid") + 
  f(ID.area.int, model="besag", graph=dat.mat, group=ID.time.int,control.group=list(model="rw2"))
mod.intIV <- inla(formula.intIV, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))
```
 
Look at the summary of the models 
```{r}

summary(mod.intI)
summary(mod.intII)
summary(mod.intIII)
summary(mod.intIV)

```
 
```{r}
dat3$rate_pos <- dat3$positive_freq/dat3$test_freq

fitI <- mod.intI$summary.fitted.values[, 1]
fitII <- mod.intII$summary.fitted.values[, 1]
fitIII <- mod.intIII$summary.fitted.values[, 1]
fitIV <- mod.intIV$summary.fitted.values[, 1]


```

Scatter plot compares the observed positivity rates and the estimated rates.
```{r}
par(mfrow=c(2,2))
plot(dat3$rate_pos, fitI, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 1 - rw2)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitII, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 2 - rw2)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitIII, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 3 - rw2)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitIV, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 4 - rw2)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)
```

Check CPO and dic, finish the code, so that we can decide on the best fitting model.
```{r}
cpo.rw2 <- c(-sum(log(mod.intI$cpo$cpo), na.rm=T), -sum(log(mod.intII$cpo$cpo), na.rm=T), -sum(log(mod.intIII$cpo$cpo), na.rm=T), -sum(log(mod.intIV$cpo$cpo), na.rm=T))
pit.rw2 <- c(-sum(log(mod.intI$cpo$pit), na.rm=T), -sum(log(mod.intII$cpo$pit), na.rm=T), -sum(log(mod.intIII$cpo$pit), na.rm=T), -sum(log(mod.intIV$cpo$pit), na.rm=T))

dic.rw2 <- c(mod.intI$dic$dic, mod.intII$dic$dic, mod.intIII$dic$dic, mod.intIV$dic$dic)
waic.rw2 <-c(mod.intI$waic$waic, mod.intII$waic$waic, mod.intIII$waic$waic, mod.intIV$waic$waic)

#sigma^2 = 1/median
sigma2.rw2 <- cbind(1/mod.intI$summary.hyperpar$`0.5quant`,
                1/mod.intII$summary.hyperpar$`0.5quant`,
                1/mod.intIII$summary.hyperpar$`0.5quant`,
                1/mod.intIV$summary.hyperpar$`0.5quant`)

tab_rw2 <- rbind(cpo=cpo.rw2, waic=waic.rw2, dic=dic.rw2, sigma2 = sigma2.rw2)
tab_rw2 <- apply(tab_rw2, 2, function(x){sprintf("%.4f",x)})
rownames(tab_rw2) <- c("CPO","WAIC","DIC","Sigma^2: spatial_unstructured",
                       "Sigma^2: spatial_structured", "Sigma^2: time_structured",
                       "Sigma^2: time_unstructured", "Sigma^2: interaction")

tab_rw2

 # inla.cpo(mod.intI,
 #          force = FALSE,
 #          verbose = TRUE,
 #          recompute.mode = TRUE)

#save(list = c("mod.intIV","mod.intI","mod.intII","mod.intIII"), file="mods_rw2_020722dat.Rdata")

#write.csv(dat3, "data_for_pred_020722.csv")
```

#### plot of random effect
```{r}
###spatial
par(mfrow=c(2,2))
spatial1 <- mod.intI$summary.random$ID.area
spatial2 <- mod.intII$summary.random$ID.area
spatial3 <- mod.intIII$summary.random$ID.area
spatial4 <- mod.intIV$summary.random$ID.area

plot(spatial1$mean ~ spatial1$ID)
plot(spatial2$mean ~ spatial2$ID)
plot(spatial3$mean ~ spatial3$ID)
plot(spatial4$mean ~ spatial4$ID)

### structured time
time_str1 <- mod.intI$summary.random$ID.time
time_str2 <- mod.intII$summary.random$ID.time
time_str3 <- mod.intIII$summary.random$ID.time
time_str4 <- mod.intIV$summary.random$ID.time

plot(time_str1$mean ~ time_str1$ID)
plot(time_str2$mean ~ time_str2$ID)
plot(time_str3$mean ~ time_str3$ID)
plot(time_str4$mean ~ time_str4$ID)

### iid time
time_unstr1 <- mod.intI$summary.random$ID.time2
time_unstr2 <- mod.intII$summary.random$ID.time2
time_unstr3 <- mod.intIII$summary.random$ID.time2
time_unstr4 <- mod.intIV$summary.random$ID.time2

plot(time_unstr1$mean ~ time_unstr1$ID)
plot(time_unstr2$mean ~ time_unstr2$ID)
plot(time_unstr3$mean ~ time_unstr3$ID)
plot(time_unstr4$mean ~ time_unstr4$ID)

```
```{r}
datout <- data.frame(dat3[1:11])
datout$estimated_pos_rate <- mod.intII$summary.fitted.values[, 1]
datout$lower_ci <- mod.intII$summary.fitted.values[, 3]
datout$upper_ci <- mod.intII$summary.fitted.values[, 5]
#write.csv(datout, "estimated_pos_rate_020722TestData_intII.csv")
```

### Models using rw1 latent effect

```{r}
#using random walk of order 1
#--- Type I interaction ---#
formula.intI.rw1 <- positive_freq ~ f(ID.area, model="bym", graph=dat.mat) + f(ID.time, model="rw1") + 
  f(ID.time2,model="iid") + f(ID.area.time, model="iid")
mod.intI.rw1 <- inla(formula.intI.rw1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type II interaction ---#
formula.intII.rw1 <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") + f(ID.time2,model="iid") +
  f(ID.area.int, model="iid", group=ID.time.int, control.group=list(model="rw1")) 
mod.intII.rw1 <- inla(formula.intII.rw1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type III interaction ---#
formula.intIII.rw1 <- positive_freq ~  f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") + f(ID.time2,model="iid") +
  f(ID.time.int,model="iid", group=ID.area.int, control.group=list(model="besag", graph=dat.mat))
mod.intIII.rw1 <- inla(formula.intIII.rw1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type IV interaction ---#
formula.intIV.rw1 <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") + f(ID.time2,model="iid") + 
  f(ID.area.int, model="besag", graph=dat.mat, group=ID.time.int,control.group=list(model="rw1"))
mod.intIV.rw1 <- inla(formula.intIV.rw1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))
```

Look at the summary of the models 
```{r}
summary(mod.intI.rw1)
summary(mod.intII.rw1)
summary(mod.intIII.rw1)
summary(mod.intIV.rw1)

```

```{r}
fitI.rw1 <- mod.intI.rw1$summary.fitted.values[, 1]
fitII.rw1 <- mod.intII.rw1$summary.fitted.values[, 1]
fitIII.rw1 <- mod.intIII.rw1$summary.fitted.values[, 1]
fitIV.rw1 <- mod.intIV.rw1$summary.fitted.values[, 1]
```

Scatter plot compares the observed positivity rates and the estimated rates.
```{r}
par(mfrow=c(2,2))
plot(dat3$rate_pos, fitI.rw1, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 1 - rw1)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitII.rw1, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 2 - rw1)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitIII.rw1, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 3 - rw1)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)

plot(dat3$rate_pos, fitIV.rw1, col=2, xlab="Observed Positivity Rate", ylab="Estimated Positivity Rate (Model 4 - rw1)", xlim=c(0, 1), ylim=c(0,1), cex.lab = 0.7)
abline(a=0, b=1, col="grey", lwd=2)
```

Check CPO and dic, finish the code, so that we can decide on the best fitting model.
```{r}
cpo.rw1 <- c(-sum(log(mod.intI.rw1$cpo$cpo), na.rm=T), -sum(log(mod.intII.rw1$cpo$cpo), na.rm=T), -sum(log(mod.intIII.rw1$cpo$cpo), na.rm=T), -sum(log(mod.intIV.rw1$cpo$cpo), na.rm=T))
dic.rw1 <- c(mod.intI.rw1$dic$dic, mod.intII.rw1$dic$dic, mod.intIII.rw1$dic$dic, mod.intIV.rw1$dic$dic)
waic.rw1 <-c(mod.intI.rw1$waic$waic, mod.intII.rw1$waic$waic, mod.intIII.rw1$waic$waic, mod.intIV.rw1$waic$waic)

#sigma^2 = 1/median
sigma2.rw1 <- cbind(1/mod.intI.rw1$summary.hyperpar$`0.5quant`,
                1/mod.intII.rw1$summary.hyperpar$`0.5quant`,
                1/mod.intIII.rw1$summary.hyperpar$`0.5quant`,
                1/mod.intIV.rw1$summary.hyperpar$`0.5quant`)

tab_rw1 <- rbind(cpo=cpo.rw1, waic=waic.rw1, dic=dic.rw1, sigma2 = sigma2.rw1)
tab_rw1 <- apply(tab_rw1, 2, function(x){sprintf("%.4f",x)})
rownames(tab_rw1) <- c("CPO","WAIC","DIC","Sigma^2: spatial_unstructured",
                       "Sigma^2: spatial_structured", "Sigma^2: time_structured",
                       "Sigma^2: time_unstructured", "Sigma^2: interaction")

tab_rw1

#higher the worse 
# mod.status <- c(mod.intI.rw1$mode$mode.status,
#                 mod.intII.rw1$mode$mode.status,
#                 mod.intIII.rw1$mode$mode.status,
#                 mod.intIV.rw1$mode$mode.status)
# mod.status

#mod.intI.rw1$logfile
```

Output estimated positivity rate
```{r}
dat3$block_grp <- as.character(dat3$block_grp)
datout <- data.frame(dat3[1:11])
datout$estimated_pos_rate <- mod.intI$summary.fitted.values[, 1]
datout$lower_ci <- mod.intI$summary.fitted.values[, 3]
datout$upper_ci <- mod.intI$summary.fitted.values[, 5]

```
### models using rw1 latent effect - without iid time

```{r}
#using random walk of order 1
#--- Type I interaction ---#
formula.intI.rw1.test <- positive_freq ~ f(ID.area, model="bym", graph=dat.mat) + f(ID.time, model="rw1")  + f(ID.area.time, model="iid")
mod.intI.rw1.test <- inla(formula.intI.rw1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type II interaction ---#
formula.intII.rw1.test <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") + 
  f(ID.area.int, model="iid", group=ID.time.int, control.group=list(model="rw1")) 
mod.intII.rw1.test <- inla(formula.intII.rw1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type III interaction ---#
formula.intIII.rw1.test <- positive_freq ~  f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") + 
  f(ID.time.int,model="iid", group=ID.area.int, control.group=list(model="besag", graph=dat.mat))
mod.intIII.rw1.test <- inla(formula.intIII.rw1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type IV interaction ---#
formula.intIV.rw1.test <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="rw1") +  
  f(ID.area.int, model="besag", graph=dat.mat, group=ID.time.int,control.group=list(model="rw1"))
mod.intIV.rw1.test <- inla(formula.intIV.rw1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))
```

```{r}
summary(mod.intI.rw1.test)
summary(mod.intII.rw1.test)
summary(mod.intIII.rw1.test)
summary(mod.intIV.rw1.test)

```
```{r}
cpo.aw1.test <- c(-sum(log(mod.intI.rw1.test$cpo$cpo), na.rm=T), -sum(log(mod.intII.rw1.test$cpo$cpo), na.rm=T), -sum(log(mod.intIII.rw1.test$cpo$cpo), na.rm=T), -sum(log(mod.intIV.rw1.test$cpo$cpo), na.rm=T))
dic.aw1.test <- c(mod.intI.rw1.test$dic$dic, mod.intII.rw1.test$dic$dic, mod.intIII.rw1.test$dic$dic, mod.intIV.rw1.test$dic$dic)
waic.aw1.test <-c(mod.intI.rw1.test$waic$waic, mod.intII.rw1.test$waic$waic, mod.intIII.rw1.test$waic$waic, mod.intIV.rw1.test$waic$waic)

cpo.aw1.test
dic.aw1.test
waic.aw1.test
```
```{r}
# save(list = c("mod.intI.rw1.test",
#               "mod.intII.rw1.test",
#               "mod.intIII.rw1.test",
#               "mod.intIV.rw1.test"), file="mod_rw1_without_timeiid.Rdata")
```

### Models using ar1 latent effect

```{r}
#using Autoregressive order 1
#--- Type I interaction ---#
formula.intI.ar1 <- positive_freq ~ f(ID.area, model="bym", graph=dat.mat) + f(ID.time, model="ar1") + 
  f(ID.time2,model="iid") + f(ID.area.time, model="iid")
mod.intI.ar1 <- inla(formula.intI.ar1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type II interaction ---#
formula.intII.ar1 <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + f(ID.time2,model="iid") +
  f(ID.area.int, model="iid", group=ID.time.int, control.group=list(model="ar1")) 
mod.intII.ar1 <- inla(formula.intII.ar1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type III interaction ---#
formula.intIII.ar1 <- positive_freq ~  f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + f(ID.time2,model="iid") +
  f(ID.time.int,model="iid", group=ID.area.int, control.group=list(model="besag", graph=dat.mat))
mod.intIII.ar1 <- inla(formula.intIII.ar1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type IV interaction ---#
formula.intIV.ar1 <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + f(ID.time2,model="iid") + 
  f(ID.area.int, model="besag", graph=dat.mat, group=ID.time.int,control.group=list(model="ar1"))
mod.intIV.ar1 <- inla(formula.intIV.ar1, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))
```

Look at the summary of the models 
```{r}
summary(mod.intI.ar1)
summary(mod.intII.ar1)
summary(mod.intIII.ar1)
summary(mod.intIV.ar1)

# save(list = c("mod.intI.ar1","mod.intII.ar1","mod.intIII.ar1","mod.intIV.ar1"), 
#      file = "ar1_model_122021_data.RData")
```
Check model
```{r}
cpo.ar1 <- c(-sum(log(mod.intI.ar1$cpo$cpo), na.rm=T), -sum(log(mod.intII.ar1$cpo$cpo), na.rm=T), -sum(log(mod.intIII.ar1$cpo$cpo), na.rm=T), -sum(log(mod.intIV.ar1$cpo$cpo), na.rm=T))
dic.ar1 <- c(mod.intI.ar1$dic$dic, mod.intII.ar1$dic$dic, mod.intIII.ar1$dic$dic, mod.intIV.ar1$dic$dic)
waic.ar1 <-c(mod.intI.ar1$waic$waic, mod.intII.ar1$waic$waic, mod.intIII.ar1$waic$waic, mod.intIV.ar1$waic$waic)

#sigma^2 = 1/median
sigma2.ar1 <- cbind(1/mod.intI.ar1$summary.hyperpar$`0.5quant`[c(1,2,3,5,6)],
                1/mod.intII.ar1$summary.hyperpar$`0.5quant`[c(1,2,3,5,6)],
                1/mod.intIII.ar1$summary.hyperpar$`0.5quant`[c(1,2,3,5,6)],
                1/mod.intIV.ar1$summary.hyperpar$`0.5quant`[c(1,2,3,5,6)])

rho_time_median <- c(mod.intI.ar1$summary.hyperpar$`0.5quant`[4],
                     mod.intII.ar1$summary.hyperpar$`0.5quant`[4],
                     mod.intIII.ar1$summary.hyperpar$`0.5quant`[4],
                     mod.intIV.ar1$summary.hyperpar$`0.5quant`[4])

group_rho_area_inte_median <- c(NA, mod.intII.ar1$summary.hyperpar$`0.5quant`[7],
                         NA, mod.intIV.ar1$summary.hyperpar$`0.5quant`[7])

tab_ar1 <- rbind(cpo=cpo.ar1, waic=waic.ar1, dic=dic.ar1, sigma2 = sigma2.ar1,
                 rho_time_median, group_rho_area_inte_median)


tab_ar1 <- apply(tab_ar1, 2, function(x){sprintf("%.4f",x)})
rownames(tab_ar1) <- c("CPO","WAIC","DIC","Sigma^2: spatial_unstructured",
                       "Sigma^2: spatial_structured", "Sigma^2: time_structured",
                       "Sigma^2: time_unstructured", "Sigma^2: interaction",
                       "rho_time_median", "group_rho_area_inte_median")

tab_ar1
```

### Models using ar1 latent effect - without iid time

```{r}
#using Autoregressive order 1
#--- Type I interaction ---#
formula.intI.ar1.test <- positive_freq ~ f(ID.area, model="bym", graph=dat.mat) + f(ID.time, model="ar1") + f(ID.area.time, model="iid")
mod.intI.ar1.test <- inla(formula.intI.ar1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type II interaction ---#
formula.intII.ar1.test <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + 
  f(ID.area.int, model="iid", group=ID.time.int, control.group=list(model="ar1")) 
mod.intII.ar1.test <- inla(formula.intII.ar1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type III interaction ---#
formula.intIII.ar1.test <- positive_freq ~  f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + 
  f(ID.time.int,model="iid", group=ID.area.int, control.group=list(model="besag", graph=dat.mat))
mod.intIII.ar1.test <- inla(formula.intIII.ar1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))

#--- Type IV interaction ---#
formula.intIV.ar1.test <- positive_freq ~ f(ID.area,model="bym",graph=dat.mat) +
  f(ID.time,model="ar1") + 
  f(ID.area.int, model="besag", graph=dat.mat, group=ID.time.int,control.group=list(model="ar1"))
mod.intIV.ar1.test <- inla(formula.intIV.ar1.test, data=dat3, 
                 family="binomial", Ntrials= new_test_freq, control.family = list(link = "logit"),  
                 control.predictor=list(compute=TRUE, link=1), 
                 control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE))
```

Look at the summary of the models 
```{r}
summary(mod.intI.ar1.test)
summary(mod.intII.ar1.test)
summary(mod.intIII.ar1.test)
summary(mod.intIV.ar1.test)
# 
#  save(list = c("mod.intI.ar1.test","mod.intII.ar1.test","mod.intIII.ar1.test","mod.intIV.ar1.test"), 
#      file = "ar1_noiidtime_model_122021_data.RData")
```

```{r}
write.csv(dat3, file = "data_for_pred.csv")
```

## Interactive map 
### Cameron county population - by count

  * CBG:480610142001 	3,816 (2010), 4054 (2019)
  * CBG:480610125042 	11,659 (2010), 14892 (2019)loaa
```{r warning=FALSE}
##extract population data
block_info <- unique(data_with_sp[,c("block_grp","census_pop")])

#merge pop data with spatial data
cameron222@data <- merge(cameron222@data, block_info, by.x="GEOID", by.y="block_grp")
cameron222@data <- cameron222@data[order(cameron222@data$ID),]

##calculate population density
##unit of ALAND is squared meters
##unit of pop_density is squared km
cameron222@data$pop_density <- cameron222@data$census_pop / as.numeric(cameron222@data$ALAND) * 1000000 

library(leaflet)
# create a interactive map
l <- leaflet() %>% addTiles()
# create the color palette (continuous)
pal <-  colorQuantile(palette = "YlOrRd", domain = cameron222$census_pop, n=5)

# the extra code building off the existing pal
pal_colors <- unique(pal(sort(cameron222$census_pop))) # hex codes
pal_labs <- quantile(cameron222$census_pop, seq(0, 1, .2)) # depends on n from pal
pal_labs <- paste(lag(pal_labs), pal_labs, sep = " - ")[-1] # first lag is NA


# add text
labels <- sprintf("<strong> BG#: %s </strong> <br/> 
                  Total population: %s <br/> 
                  Population density (per squared km): %.2f",
                  cameron222$GEOID, cameron222$census_pop, cameron222$pop_density) %>%
  lapply(htmltools::HTML)

# add labels to the interactive map
interactive <-  l %>%
  addPolygons(
    group="marker",
    color = "grey", 
    weight = 1,
    fillColor = ~ pal($census_pop), 
    fillOpacity = 0.5,
    highlightOptions = highlightOptions(weight = 4),
    label = labels,
    labelOptions = labelOptions(
      style = list( "font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px", direction = "auto")
  ) %>%
  addLegend(
    colors=pal_colors, labels=pal_labs, opacity = 0.5,
    title = "Cameron county population", position = "bottomright"
  )


options(digits = 3)



```

### Cameron county population - by population density

```{r}
# create a interactive map
l2 <- leaflet() %>% addTiles()
# create the color palette (continuous)
pal2 <-  colorQuantile(palette = "YlOrRd", domain = cameron222$pop_density, n=5)

# the extra code building off the existing pal
pal_colors2 <- unique(pal2(sort(cameron222$pop_density))) # hex codes
pal_labs2 <- sprintf("%.2f", quantile(cameron222$pop_density, seq(0, 1, .2))) # depends on n from pal
pal_labs2 <- paste(lag(pal_labs2), pal_labs2, sep = " - ")[-1] # first lag is NA


# add text
labels <- sprintf("<strong> BG#: %s </strong> <br/> 
                  Total population: %s <br/> 
                  Population density (per squared km): %.2f",
                  cameron222$GEOID, cameron222$census_pop, cameron222$pop_density) %>%
  lapply(htmltools::HTML)

# add labels to the interactive map
interactive2 <-  l2 %>%
  addPolygons(
    group="marker",
    color = "grey", 
    weight = 1,
    fillColor = ~ pal2($pop_density), 
    fillOpacity = 0.5,
    highlightOptions = highlightOptions(weight = 4),
    label = labels,
    labelOptions = labelOptions(
      style = list( "font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px", direction = "auto")
  ) %>%
  addLegend(
    colors=pal_colors2, labels=pal_labs2, opacity = 0.5,
    title = "Cameron county population", position = "bottomright"
  )


options(digits = 3)
```

**Map of Cameron county population - by count**

<br>
```{r, echo=FALSE, fig.width=10, fig.height=8}
interactive
```

**Map of Cameron county population - by population density**

<br>
```{r, echo=FALSE, fig.width=10, fig.height=8}
interactive2
```



## Comparing positivity rate observed vs estimated
```{r echo=FALSE}
#######function of comparing positivity rate
dat4 <- data.frame(data_with_sp[1:11])
dat4$estimated_pos_rate <- mod.intI$summary.fitted.value[,1]
dat4$lower_ci <- mod.intI$summary.fitted.value[,3]
dat4$upper_ci <- mod.intI$summary.fitted.value[,5]

  
library(lubridate)
all_week <- ymd( "2020-03-16" ) + weeks(1:max(data_block_gap$week) - 1 )
my_lab <- NULL
n <- 1
for (i in 1:(92%/%4)) {
  n <- 1 + 4*(i-1)
  my_lab <- c(my_lab, all_week[n])
}
my_lab <- as.Date(my_lab, origin = "1970-01-01")

plot_pos_rate <- function(cbg_id){
    grp <- dat4[dat4$block_grp == cbg_id,]
    
    ggplot(grp, aes(x = week)) + 
              geom_line(aes(y = estimated_pos_rate, color = "Estimated")) + 
              geom_line(aes(y = positive_rate, color = "Observed")) +
              geom_ribbon(aes(ymin=lower_ci, ymax=upper_ci), linetype=2, alpha = 0.4)+
              ylab("Positivity rate") +   
              theme(strip.background =element_rect(fill="white")) +
              theme_light() +
              scale_color_manual("Positivity rate", values = c("Estimated" = "red", "Observed" = "blue")) +
              ggtitle(paste0("Comparing positivity rate"," - ",cbg_id)) + 
              scale_x_continuous(breaks = seq(1,92, by = 4), labels = my_lab) +
              theme(axis.text.x = element_text(angle = 75, vjust = 0.5, size = 5))
}


plot_pos_rate("480610136001")#extreme
plot_pos_rate("480610120014")#extreme
plot_pos_rate("480610125042")
plot_pos_rate("480610144001")
plot_pos_rate("480610132051")#random select
plot_pos_rate("480610133031")#random select

plot_pos_rate("480610122005")#contain positivity rate > 1

##output estimated positivity rate
#write.csv(dat2, "estimated_positivity_rate.csv")
```

Distribution of positivity rate
```{r eval=FALSE, include=FALSE}
hist(dat2$positive_rate, main = "Observed positivity rate")
hist(model_logit$summary.fitted.value[,1], main = "Fitted positivity rate")
```